---
params:
  country: "Greece"
  year: 2022
  month: 2
---
# COVID-19 deaths and cases parameterized

This report visualizes the number of COVID-19 cases and deaths in a selected European country for a specific month and year, based on parameters defined. By adjusting the country, year, and month parameters, the same report can be reused to explore trends in different regions and time periods.


## Obtaining the data and loading the necessary libraries

The data used in the graphs is obtained from the *European Centre for Disease Prevention and Control* under the dataset titled *Data on the daily number of new reported COVID-19 cases and deaths by EU/EEA country*. Please note that the data has not been updated since November 1, 2022.

```{r imports, warning= FALSE, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(plotly)

data <- read.csv("https://opendata.ecdc.europa.eu/covid19/nationalcasedeath_eueea_daily_ei/csv", na.strings = "", fileEncoding = "UTF-8-BOM")

```


## Preparing the data

The data is filtered by the parameters chosen by the viewer of the plots.

```{r filtering}

#use paramaters from yaml header
filtered_data <- data %>%
  filter(countriesAndTerritories == params$country, year == params$year, month == params$month)

# Fix date format for ggplot
filtered_data$dateRep <- as.Date(filtered_data$dateRep, format = "%d/%m/%Y")


```

## Plotting the COVID-19 death counts by parameters

```{r plot death}

# Plot the death count
deaths <- filtered_data %>% 
  ggplot(aes(x = dateRep, y = deaths)) + 
  geom_line(color = "darkred", linewidth= 1.5 ) + 
  theme_classic() + 
  labs(
    title = paste("COVID-19 Deaths in", params$country, "(", params$month, "/", params$year, ")"),
    x = "Date",
    y = "Number of Deaths"
  ) + #Rotating the axis a little for better visualization
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplotly(deaths)

```

## Plotting the COVID-19 cases by parameters*

```{r plot cases}

# Plot the case count
cases <- filtered_data %>% 
  ggplot(aes(x = dateRep, y = cases)) + 
  geom_line(color = "orange", linewidth = 1.5) + 
  theme_bw() + 
  labs(
    title = paste("COVID-19 cases in", params$country, "(", params$month, "/", params$year, ")"),
    x = "Date",
    y = "Number of cases"
  ) + #Rotating the axis a little for better visualization
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplotly(cases)

```
